Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})

### 2014-05-19 manuscript.tex
edits to discussion

\jri{do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}

The phenotypic effects of the \emph{Hopscotch} insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}. Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion in individuals sampled from natural populations of teosinte. \jri{isn't this what weber did?? for 70+ populations!?} However, we found no significant effect of the \emph{Hopscotch} on tillering index or tiller number in our phenotyping experiments, and the effect of the \emph{Hopscotch} insertion in teosinte is discordant with that of maize. The lack of correlation between \emph{Hopscotch} genotype and tillering index or tiller number is surprising given its effects in maize. It is certainly possible that even though previous data demonstrate an effect of the \emph{Hopscotch} on tillering in maize \citep{Studer2011}, that the effect of the \emph{Hopscotch} in teosinte is more complicated and may be more difficult to observe. Moreover, \emph{tb1} is a single gene in a complex pathway that affects branching and tillering traits, and perhaps in combination with alleles at other loci the phenotypic effects of the \emph{Hopscotch} on tillering may not be consistent. \jri{thsi section still needs work. i think we have to do more here. weber shows an association between SNPs in tb1 and branch length. we need to discuss that!}

>>>>

\jri{do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check} \mbh{I've added a few sentences on selection.  Do we still want to compare sequences with and without \emph{Hopscotch}?  I agree its a good idea and could end up being really interesting.  Perhaps something we could look at after submission and incorporate during revisions?}

Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Clark2006, Studer2011}.  \citet{Weber2007} have described significant phenotypic associations between genic markers in \emph{tb1} and lateral branch length and female ear length within a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}. Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte. We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number in our phenotyping experiments, a result that is discordant with its clear phenotypic effects in maize. One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci. In fact, \emph{tb1} is known to be a single gene in a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007}.
